### 2013-06-28 inst/doc/reml.Rmd
draft manuscript

* The Ecological Metadata Language (EML) has been around for a while now, but has so far seen limited use.

* R is a widely adopted among the community, many researchers already familiar with reading and writing data from R.

* R provides a rich set of data types and constructs that can be mapped into EML (e.g. integers, floats, ordered/nominal factors, character strings)

### Current state of EML

* Generation of EML is limited.  e.g. to LTER centers with informatics support. 
* Applications of existing EML more limited still (?)

* Software to work with EML is limited to tools with low adoption rates across the community (e.g. Kepler).

 Researchers are isolated from the metadata encoding, instead providing plain-text descriptions in MS Word documents.  Allows for low barrier of entry but also makes a barrier to richer encoding.

### Promise of REML

### Authoring tools

### Parsing tools

### Publishing tools