### 2013-06-29 inst/doc/reml.Rmd
documentation

## Why EML?

Archival, open data is great ... blah blah.  Good reuse is all about good Metadata blah blah...

But specifying metadata is typically onerous, time-consuming and error prone.  Your metadata doesn't define your date format/missing value character/etc etc?  What, that's metadata too?  Wouldn't it be nice if someone thought of all the things we should record and just made a template we could use?

Oh, wait, they did.  It's called EML, and it's been around a long time.  Like carved in petroglyphs long time, in the days before a public internet existed.

Oh, you haven't heard of it?  That's probably because what should be a time-saving template appears more like a huge technological barrier best left to tech geeks.   But it isn't.

> Terence Parr once said "being an expert in XML is like being an expert in CSV".

Or at least, it doesn't have to be that way.  We just need better tools.

## Why REML?

* The Ecological Metadata Language (EML) has been around for a while now, but has so far seen limited use.

* R is a widely adopted among the community, many researchers already familiar with reading and writing data from R.

* R provides a rich set of data types and constructs that can be mapped into EML (e.g. integers, floats, ordered/nominal factors, character strings)

### Current state of EML

* Generation of EML is limited.  e.g. to LTER centers with informatics support. 
* Applications of existing EML more limited still (?)

* Software to work with EML is limited to tools with low adoption rates across the community (e.g. Kepler).

 Researchers are isolated from the metadata encoding, instead providing plain-text descriptions in MS Word documents.  Allows for low barrier of entry but also makes a barrier to richer encoding.

### Promise of REML

### Authoring tools

### Parsing tools

### Publishing tools