### 2013-07-07 inst/doc/reml-blog-post.Rmd
documentation updated

You know the spiel.  Archivally preserved open data is great [1]. For you [2], for your citations [3], for your grants, [4] and of course, for scientific progress more generally [5].

And you know that making data reusable is all really all about having good metadata [6-398].

But specifying metadata is typically onerous, time-consuming and error prone.  Your metadata doesn't define your date format/missing value character/etc etc?  What, that's metadata too?  Wouldn't it be nice if someone thought of all the things we should record and just made a template we could use?

Oh, wait, they did.  It's called EML, and it's been around a long time.  Like carved in petroglyphs long time, in the days before a public internet existed.

Oh, you haven't heard of it?  That's probably because what should be a time-saving template appears more like a huge technological barrier best left to tech geeks.   But it isn't.

Or at least, it doesn't have to be that way.  We just need better tools.

You already use XML every day.  Microsoft office?  Those `.docx`, `xlsx`, etc., are just a bunch of XML files zipped together (just try `unzip file.docx`)  The tools are just good enough to hide the XML from you.  XML is so general that it's essentially meaningless to say a file is XML.  Could be anything. Terence Parr once said "being an expert in XML is like being an expert in CSV".

That's where schema come in.

## Why REML?

* The Ecological Metadata Language (EML) has been around for a while now, but has so far seen limited use.

* R is a widely adopted among the community, many researchers already familiar with reading and writing data from R.

* R provides a rich set of data types and constructs that can be mapped into EML (e.g. integers, floats, ordered/nominal factors, character strings)