### 2013-07-07 inst/doc/reml.Rmd

Abstract

1. Recording good metadata is at the heart of effective data synthesis and reuse
2.

Vertically integrated repositories vs metadata driven repositories.

Data synthesis

## Why EML?

But specifying metadata is typically onerous, time-consuming and error prone.

Such definitions are also not machine readable.

Or at least, it doesn't have to be that way.  We just need better tools.

## Why REML?

* The Ecological Metadata Language (EML) has been around for a while now, but has so far seen limited use.

* R is a widely adopted among the community, many researchers already familiar with reading and writing data from R.

* R provides a rich set of data types and constructs that can be mapped into EML (e.g. integers, floats, ordered/nominal factors, character strings)

### Current state of EML

* Generation of EML is limited.  e.g. to LTER centers with informatics support. 
* Applications of existing EML more limited still (?)

* Software to work with EML is limited to tools with low adoption rates across the community (e.g. Kepler).

 Researchers are isolated from the metadata encoding, instead providing plain-text descriptions in MS Word documents.  Allows for low barrier of entry but also makes a barrier to richer encoding.

### Promise of REML

### Getting Started: A minimal example.

Imagine we are working with a `data.frame` in R analgous to Table 1.

<!--Writing a data.frame into EML -->

Writing this table into EML

### Authoring tools

### Parsing tools

### Publishing tools