### 2013-08-04 inst/examples/S4_draft_methods.Rmd
examples of what S4 coercion methods might look like; still working out questions

## What example functions would look like
```{r}
eml_dataset <- function(x)
  as(x, "eml:dataset")

eml_write <- function(x, file = NULL){
  eml <- as(x, "eml:eml")
  doc <- newXMLDoc(node=eml)
  saveXML(doc, file=file)
}
```

some testing data

```{r}
dat = data.frame(river = c("SAC",  "SAC",   "AM"),
                 spp   = c("king",  "king", "ccho"),
                 stg   = c("smolt", "parr", "smolt"),
                 ct    = c(293L,    410L,    210L))
```

New metadata format

```{r}
metadata <- 
  list("river" = list("river",
                      "River site used for collection",
                      c(SAC = "The Sacramento River", AM = "The American River")),
       "spp" = list("spp",
                    "Species common name", 
                    c(king = "King Salmon", ccho = "Coho Salmon")),
       "stg" = list("stg",
                    "Life Stage", 
                    c(parr = "third life stage", smolt = "fourth life stage")),
       "ct"  = list("ct",
                    "count", 
                    "number"))
```

### Coercion methods

Take this metadata and coerce it into a valid `attributeList`
A list of attributes based on this structure would look like:

```{r}

  setAs("list", "eml:attribute", 
        function(from){
         new("eml:attribute", 
             attributeName = from[[1]],
             attributeDefinition = from[[2]], 
             measurementScale = as(from[[3]], from[[4]])) 
        })
```

The trick is to decide what class measurementScale should be coerced into.  This should come from the data.frame class. Not clear if that's allowed in a standard coercion.  We could:

- [ ] Define the `setAs` method inside a function taking both the metadata list and the data table.

```{r}
detect_class <- function(dat, meta){
  for(i in length(dat)){
    meta[[i]][[4]] = map(class(dat[[i]])) 
  }
  meta 
}

map <- function(x){
  if(is(x, "numeric"))
    "eml:ratio"
  else if(is(x, "ordered"))
    "eml:ordinal"
  else if(is(x, "nominal"))
    "eml:nominal",
  else if(is(x, "POSIXlt") | is(x, "POSIXct"))
    "eml:dateTime"
  else if(is(x, "character"))
    "eml:nominal"
}
```

Then in any function that gets both dat and metadata, we can correct things:

```{r}
metadata <- detect_class(dat, metadata)
```
- [ ] Something else?

If/once we know it's a nominal, conversion is easy:

```{r}
setAs("character", "eml:nominal", 
      function(from)
        new("eml:nominal",
            nonNumericDomain = as(from, "eml:nonNumericDomain")))

setAs("character", "eml:textDomain", function(from){
  new("eml:textDomain", definition = from)
})

setAs("character", "eml:nonNumericDomain", 
      function(from){
        if (length(from) == 1)
          new("eml:nonNumericDomain",
              textDomain = as(from, "eml:textDomain"))
        else if(length(from) > 1)
          new("eml:nonNumericDomain", 
            enumeratedDomain = as(from, "ListOfCodeDefinition"))
      })

setAs("character", "ListOfCodeDefinition", 
      function(from){
        new("ListOfCodeDefinition", 
            lapply(names(from), function(name)
               new("eml:codeDefinition", code = name, definition = from[name])))
      })
```

Then convert the "unit metadata" for the river values just by:

```{r}
as(metadata$river[[3]], "nominal")
```

`nominal` is potential slot for `measurementScale`, which is a slot of an `attribute`.

Moving up the DOM tree from here is easy:

```{r}
attrList = new("eml:attributeList", as(metadata_list, "eml:attribute")
new("eml:dataSet", dataTable = new("eml:dataTable", attributeList = attrList) )
```

and so forth.  lots of permutations are possible.  We can define coercion methods and merge methods to make this all smooth and flexible.

## We either want to look at the data.frame column classes,
## Or guess from the length of the metadata

setAs("character", "eml:measurementScale",
      function(from){
        if(length(from) > 1) ## assume factor.  Not very robust!
          new("eml:measurementScale", 
              nominal = as(from, "eml:nominal"))
        else if(length(from) == 1) ## Assume a unit of 
      })