### 2013-09-03 inst/doc/vingette.Rmd
updated README

reml
====

An R package for reading, writing, integrating and publishing data using the Ecological Metadata Language (EML) format.

* **Note:** This package is still in active development and not yet submitted to CRAN.  Functions and documentation may be incomplete and subject to change.  
* Maintainer: Carl Boettiger
* Authors: Carl Boettiger, Duncan Temple Lang, Karthik Ram, Matt Jones
* License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
* [Issues](https://github.com/ropensci/reml/issues): Bug reports, feature requests, and development discussion.

_This is a dynamic document, generated by [knitr](http://yihui.name/knitr) using [vignette.Rmd](https://github.com/ropensci/reml/blob/master/inst/doc/vingette.Rmd) as the source._

```{r include=FALSE, cache=FALSE}
source("~/.knitr_defaults.R")
opts_chunk$set(cache=FALSE)
```

Quick Start
===========

Install the R package:

```{r, results="hide", message=FALSE, warning=FALSE, eval=FALSE}
library("devtools")
install_github("reml", "ropensci")
```

Load the package:

```{r, results="hide", message=FALSE, warning=FALSE}
library("reml")
```

Writing R data into EML
-----------------------

Consider some dataset as an R `data.frame`.

```{r}
dat = data.frame(river = c("SAC", "SAC", "AM"),
                 spp   = c("king", "king", "ccho"),
                 stg   = c("smolt", "parr", "smolt"),
                 ct    =  c(293L, 410L, 210L))
```

Provide a list giving the  __column header__ used in the data.frame, followed by a plain text definitions for the column, followed by a character vector giving the definition for the units:

__column descriptions__: These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful. See "Advanced Use" for details on adding richer information, such as the method used to collect the data or set the geographic, taxanomic, or temporal coverage of an individual column.

__column units__:   For factors, this is a definition of the levels involved.  For numeric data, specify the units from [this list](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-unitTypeDefinitions.html#StandardUnitDictionary).  For dates, specify the format, (e.g. YYYY or MM-DD-YY). For character strings, a definition of the kind of string can be given, (e.g. species scientific name), otherwise the column description will be used.

```{r}
metadata <- 
  list("river" = list("river",
                      "River site used for collection",
                      c(SAC = "The Sacramento River", 
                        AM = "The American River")),
       "spp" = list("spp",
                    "Species common name", 
                    c(king = "King Salmon", 
                      ccho = "Coho Salmon")),
       "stg" = list("stg",
                    "Life Stage", 
                    c(parr = "third life stage", 
                      smolt = "fourth life stage")),
       "ct"  = list("ct",
                    "count", 
                    "number"))

```

*for convenience, this metadata list can instead be constructed with the help of the reml wizard*.  Just use `metadata <- metadata_wizard(dat)` to begin. While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.

```{r}
eml_write(dat, metadata, title = "reml example",  
          description = "An example, intended for
                              illustrative purposes only.",
          creator = "Carl Boettiger <cboettig@gmail.com>",
          file = "reml_example.xml")
```

See the [EML generated](https://github.com/ropensci/reml/tree/master/inst/doc/reml_example.xml) by this example.  Note that if the `metadata` argument providing the definitions for the `data.frame` is not specified, `reml` will launch the wizard to prompt the user.

### Configuration

Configure general metadata you may want to frequently reuse, avoiding having to specify things like the name of the creator or contact.

```{r}
eml_config(creator = list("Carl Boettiger <cboettig@ropensci.org>", "Karthik Ram"), contact = "Carl Boettiger <cboettig@ropensci.org>")
```

* See [Open issues in writing EML](https://github.com/ropensci/reml/issues?labels=write+eml&page=1&state=open)

Publish EML
-----------

Once we have generated a basic EML file, we can publish it to a variety
of academic repositories where it will be indexed, permenently archived,
more easily citable, and publicly discoverable.

In this example, we will publish the above
EML to [figshare](http://figshare.com) using the
[rfigshare](https://github.com/ropensci/rfigshare) package. Like most
academic repostories, figshare requires some basic metadata for any entry.
Much of this we can automatically extract from our existing EML file. 
Additional metadata unique to figshare (e.g. matching it's allowed "categories"
can also be provided.

_Note: this example assumes `rfigshare` is installed, even though it is not 
required to use reml and run the examples in the other sections_

<!-- We don't want to generate a DOI every time we run the vignette -->
```{r publish_figshare} 
id = eml_publish("reml_example.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML", destination="figshare")
```

This creates a draft file visible only to the user configured in `rfigshare`.  The document can be made (permanently) public using either the figshare web interface, the `rfigshare` package (using `fs_make_public(id)`) or just by adding the argument `visibility = TRUE` to the above `eml_publish` command.

<!-- In return, figshare provides the object with a DOI, which is added to the EML.

```{r figshare_doi, eval=FALSE}
doc <- eml_read("my_eml_data.xml")
citation(doc) # not implemented yet
```
-->

<-- Too detailed for this vignette

For the sake of completeness, the explicitly figshare metadata is also written into a 
dedicated `<additionalMetadata>` node with the attribute `id = figshare`.  This allows
us to distinguish between metadata that is an is not available to figshare's indexing,
e.g. if we use the figshare search tool.

-->

#### In development

See [Open issues in publishing EML](https://github.com/ropensci/reml/issues?labels=publish+eml&page=1&state=open)

Reading EML
-----------

#### In development

See [Open issues in reading EML](https://github.com/ropensci/reml/issues?labels=read+eml&page=1&state=open)

Integrating multiple EML files into a single data frame
-------------------------------------------------------

#### In development

See [Open issues in integrating EML](https://github.com/ropensci/reml/issues?labels=integrate+eml&page=1&state=open)

Testing and development
-----------------------

While the dependencies for basic functionality are kept to a minimum, to access all the functions and tests implemented in `reml` you'll need several additional packages.

```r
install.packages(c("yaml", "knitr", "rfigshare", "testthat", "RCurl"))
```

Some of these additional packages are not yet on CRAN and may not be stable.

```r
install.packages("uuid",,'http://rforge.net/',type='source')
install_github("RHTMLForms", "omegahat")
install_github("XMLSchema", "omegahat")
```

Successful installation of these packages will require the `devtools` package (for `install_github`) and the ability to build packages from source.