### 2014-01-21 inst/doc/README.Rmd
rename package as EML instead of reml, closes #63

EML
====

An R package for reading, writing, integrating and publishing data using the Ecological Metadata Language (EML) format.

* **Note:** This package is still in active development and not yet submitted to CRAN.  Functions and documentation may be incomplete and subject to change.  
* Maintainer: Carl Boettiger
* Authors: Carl Boettiger, Claas-Thido Pfaff, Duncan Temple Lang, Karthik Ram, Matt Jones
* License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
* [Issues](https://github.com/ropensci/EML/issues): Bug reports, feature requests, and development discussion.

_This is a dynamic document, generated by [knitr](http://yihui.name/knitr) using [vignette.Rmd](https://github.com/ropensci/EML/blob/master/inst/doc/vingette.Rmd) as the source._

```{r include=FALSE, cache=FALSE}
source("~/.knitr_defaults.R")
```

Quick Start
===========

Install the R package:

```{r, results="hide", message=FALSE, warning=FALSE, eval=FALSE}
library("devtools")
install_github("EML", "ropensci")
```

Load the package:

```{r, results="hide", message=FALSE, warning=FALSE}
library("EML")
```

Writing R data into EML
-----------------------

`EML` now extends R's `data.frame` class by introducing the `data.set` class which includes additional metadata required by EML.  A `data.set` can be created much like a `data.frame` by specifying additional arguments

```{r}
dat = data.set(river = c("SAC",  "SAC",   "AM"),
               spp   = c("king",  "king", "ccho"),
               stg   = c("smolt", "parr", "smolt"),
               ct    = c(293L,    410L,    210L),
               col.defs = c("River site used for collection",
                            "Species common name",
                            "Life Stage", 
                            "count of live fish in traps"),
               unit.defs = list(c(SAC = "The Sacramento River", 
                                  AM = "The American River"),
                                c(king = "King Salmon", 
                                  ccho = "Coho Salmon"),
                                c(parr = "third life stage", 
                                  smolt = "fourth life stage"),
                                "number"))

```

- `col.defs`: These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful. See "Advanced Use" for details on adding richer information, such as the method used to collect the data or set the geographic, taxanomic, or temporal coverage of an individual column.

- `unit.defs`:   For factors, this is a definition of the levels involved.  For numeric data, specify the units from [this list](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-unitTypeDefinitions.html#StandardUnitDictionary).  For dates, specify the format, (e.g. YYYY or MM-DD-YY). For character strings, a definition of the kind of string can be given, (e.g. species scientific name), otherwise the column description will be used.

Alternatively, annotations can be added to an existing data frame, `data.set(my.data.frame, col.defs = ..., unit.defs = ...)`.

We will also specify a default creator who will also be used as the contact person for EML files created in this session.

```{r} 
eml_config(creator="Carl Boettiger <cboettig@gmail.com>")
```

While we will always be able to provide alternative or additional creators or contact person later, `eml_config` can store the defaults to save us the trouble of entering this data repeatedly.  See `?eml_config` for more information and other configurable defaults.

With this information in place, we have all the required metadata to generate a minimally valid EML file documenting this dataset.

```{r}
eml_write(dat, file = "EML_example.xml")
```

*for convenience, had we ommitted any essential metadata, such as providing only an unannotated `data.frame` in place of a `data.set`, `EML` will launch its metadata wizard to assist in constructing the metadata based on the data.frame provided*.  While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.

See the [EML generated](https://github.com/ropensci/EML/tree/master/inst/doc/EML_example.xml) by this example.

We can validate the resulting document to check that no required fields are missing and that no entry has been incorrectly formatted.  This ensures that other software can correctly parse and interpret the metadata document:

```{r}
eml_validate("EML_example.xml")
```

* See [Open issues in writing EML](https://github.com/ropensci/EML/issues?labels=write+eml&page=1&state=open)

Publish EML
-----------

Once we have generated a basic EML file, we can publish it to a variety
of academic repositories where it will be indexed, permenently archived,
more easily citable, and publicly discoverable.

### Publishing to figshare

In this example, we will publish the above
EML to [figshare](http://figshare.com) using the
[rfigshare](https://github.com/ropensci/rfigshare) package. Like most
academic repostories, figshare requires some basic metadata for any entry.
Much of this we can automatically extract from our existing EML file. 
Additional metadata unique to figshare (e.g. matching it's allowed "categories"
can also be provided.

_Note: this example assumes `rfigshare` is installed, even though it is not 
required to use EML and run the examples in the other sections.  See the 
`rfigshare` documentation for authenticating `rfigshare` with your user account_

<!-- We don't want to generate a DOI every time we run the vignette -->

```{r publish_figshare} 
eml_publish("EML_example.xml", description="Example EML file from EML", categories = "Ecology", tags = "EML", destination="figshare")
```

This creates a draft file visible only to the user configured in
`rfigshare`.  The document can be made (permanently) public using
either the figshare web interface, the `rfigshare` package (using
`fs_make_public(id)`) or just by adding the argument `visibility = TRUE`
to the above `eml_publish` command.

#### In development

See [Open issues in publishing EML](https://github.com/ropensci/EML/issues?labels=publish+eml&page=1&state=open)

Parsing and exploring an EML file
----------------------------------

Users will not only want to generate EML files, but also take advantage of
the wealth of data already well documented in EML.  The package provides
many tools to facilitate the extraction of data from EML files into
their native R formats, as well as utilities to access and manipulate
the metadata provided. By having these tools avialble in the scripted R
environment, it becomes possible to write functions that can be easily
applied across large collections of files. In this way, researchers
can scale their analyses across ever larger collections of datasets by
automating the more tedious aspects of data discovery and integration.

```{r}
obj <- eml_read("EML_example.xml")
```

We can also read in a remote file by providing a URL or KNB object identifier (such as a DOI).

The `eml_get` function provides us with easy access to many of the component elements of the metadata file.  See the documentation for a complete list.

```{r}
dat <- eml_get(obj, "data.set")
```

```{r}
eml_get(obj, "contact")
```

```{r}
eml_get(obj, "citation_info")
```

Advanced writing of EML
-----------------------

**IN PROGRESS**

Our minimal EML file barely scratches the surface of what is possible to do with EML.  In [Advanced writing of EML](), we construct a much richer EML file, including:

- Constructing more complete lists of authors, publishers and contact.
- Summarizing the geographic, temporal, and taxonomic coverage of the dataset
- Reading in pages of methods descriptions from a Word document
- Adding arbitrary additional metadata
- Indicating the canonical citation to the paper that should be acknowledged when the data is re-used.  
- Conversion between EML and other metadata formats, such as NCBII and ISO standards.

In so doing we will take a more modular approach that will allow us to build up our metadata from reusable components, while also providing a more fine-grained control over the resulting output fields and files.

Advanced parsing and manipulation of EML 
----------------------------------------

**IN PROGRESS**

In [Advanced parsing and manipulation of EML](), we

- Introduce how to access any EML element in R using the S4 subsetting mechanism
- Demonstrate how to extract and manipulate semantic RDF triples from EML metadata
- Illustrate how EML metadata extraction can be applied across a collection of files 
- Illustrate how EML manipulation can be combined with queries across even larger distributed collections of EML data over the dataone network
- Demonstrate merging of multiple datasets based on semantics

Testing and development
-----------------------

While the dependencies for basic functionality are kept to a minimum, to access all the functions and tests implemented in `EML` you'll need several additional packages.

```r
install.packages(c("yaml", "knitr", "rfigshare", "testthat", "RCurl"))
```

Some of these additional packages are not yet on CRAN and may not be stable.

```r
install.packages("uuid")
install_github("RHTMLForms", "omegahat")
install_github("XMLSchema", "omegahat")
```

Successful installation of these packages will require the `devtools` package (for `install_github`) and the ability to build packages from source.

***Suggestions, feature requests, bug reports and other feedback welcome!*** Please see our [issues tracker](https://github.com/ropensci/EML/issues)