### 2014-04-22 inst/doc/pubs/manuscript.Rmd
update affiliations

---
layout: preprint
#layout: review, 11pt
title: "Ecological Metadata Language interface for R: facilitating synthesis and integration of heterogenous data"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig@gmail.com
    footnote: Corresponding author
  - name: Karthik Ram 
    affiliation: Berkeley 
  - name: Class-Thido Pfaff
    affiliation: Leipzig
  - name: Matt Jones 
    affiliation: NCEAS 
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: Berkeley
    address: |
      Department of Environmental Science and Policy Management,
      University of California, Berkeley, 130 Mulford Hall,
      Berkeley, CA 94720. USA
  - code: Leipzig
    address: 
  - code: NCEAS
    address |
      National Center for Ecological Analysis and Synthesis,
      735 State Street, Suite 300, 
      Santa Barbara, CA 93101

abstract: | 
  1. _Write something about the importance of data, data publication, sharing
     and annotation_

  2. _Write something about EML_

  3. The EML package provides automated and scriptable tools to make it
     easy to generate and publish data and metadata using state-of-the-art
     technology and publishing standards.

  4. The EML package also makes it easy to consume richly structured data
     and metadata from rapidly growing archives of repositories such as
     KNB and DataONE, including over 20,000 datasets deposited by Long
     Term Ecological Research (LTER) areas over the past 30 years

  5. The EML package facilitates manipulation and integration of
     heterogeneous data sources by providing simple functions to add rich,
     machine-readable annotations to almost any data set.  These
     descriptions can then be used to enable programmatic recombination and
     synthesis of different data sets, allowing users to both take better
     advantage of their own data collections and more easily integrate
     them with others.

---

<!--
The future belongs to those who can leverage existing data and provide reusable data.  
-->

Heterogeneous Data
[Jones 2006], [Green 2005]

Data in ecology and evolutionary biology research collected for a
particular purpose can often prove valuable to an entirely different
line of research. One study may collect morphometric measurements of
coral reef fish taken to understand their evolution, while another
reuses only the species names and geographic locations of the samples,
along with other data, to aid the design of marine reserves.  Such reuse
is fundamental to much of our research, and yet it is often difficult
or laborious to discover and obtain. The publications are the only
documentation of what data is available.  The distribution and discovery
of publications is geared towards reaching audience interested in common
set of questions, not a common set of data. Even within a subdispline,
re-use of data is often laborious or impossible due to lack of data or
sufficient documentation of the data.

Vertically intergrated data repositories such as Genbank, GBIF [Canhos 2004],
or TreeBase [Morell 1996] provide one route to addressing the challenges 
of data integration.  These data warehouses provide all there data in a common 
structure that can usually be queried through a web interface. This makes it
relatively easy to discover available data and access it in a consistent and
predictable format.  However, such utility comes at a cost.  This approach
discards the original data model, preserving only the data that can be shoehorned
into a common data standard.

<!-- is "data model" more confusing than "data format"? -->

These challenges have long been recognized and broadly accepted that
they can be best addressed through establishing and using appropriate
metadata [Michener 1997]. In this approach it is the metadata, rather than
the data itself, that conforms to a standard.  The Knowledge Network for
Biocomplexity (KNB) is one example of a metadata-based ecological data 
repository [Jones 2001], [KNB].  The metadata approach preserves the original data
model from the collecting study. Rather than requiring that the raw data
be preserved in a standard format -- standard file types, standard column headings, 
standard date formats, missing value formats, etc etc, -- the metadata standard
approach provides a consisent way to document whatever data model the researchers
use.  In principle, this approach is both more powerful than vertical integration, 
(as the full context of the data is preserved), and easier to implement for both
researcher and repository (as the researcher does not to to convert data to 
the format required by the repository, and very heterogenous data models can be 
stored and searched for in a single system).   [Jones 2006] provides
an excellent explanation and comparison of these approaches.

Unfortunately, adoption of these metadata standards for annotating research
data has been limited, owing to both technological and social challenges 
involved [Reichman 2011].

Things to discuss:

*  It's not just a problem of sharing culture 
* Sharing is not effective: close competitors can still scoop you, but broader reach can't be realized without good metadata 
* [Strasser] survey showing lagging data management practices
* [Stodden 2010] survey showing effort in curating data is a barrier

Consequently ecological
data reuse remains difficult and haphazard.

Widely adopted computing environments such as the R statistical sofware
provide a key opportunity to increase adoption of metadata standards.

Methods in Ecology and Evolution has published XXX application notes or articles presenting new software packages.  XXX of them are R packages.

[Reichman 2011]: http://dx.doi.org/10.1126/science.1197962
[Jones 2006]: http://dx.doi.org/10.1146/annurev.ecolsys.37.091305.110031
[KNB]: http://knb.ecoinformatics.org
[Jones 2001]: http://dx.doi.org/10.1109/4236.957896
[Michener 1997]: http://dx.doi.org/10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2
[Morell 1996]: http://dx.doi.org/10.1126/science.273.5275.569
[Canhos 2004]: https://journals.ku.edu/index.php/jbi/article/viewArticle/3 
[Michener 2006]: http://dx.doi.org/10.1016/j.ecoinf.2005.08.004
[Michener 2012]: http://dx.doi.org/10.1016/j.tree.2011.11.016

1. Commonly used data structures in R provide basic metadata automatically
2. We can more easily avoid redundant metadata entry 
3. Automate quality control checks
4. Publish
5. Arbitrarily complex search 
6. Building a synthesis platform

### Automatic metadata detection

Not sure this counts as a section - but we can automatically detect
at least the difference between numeric and character data, factors,
ordered factors, etc.

### Reusing fields

- Examples of filling in contact information, etc using `eml$set`,
using an existing EML file as a template, and using an external yaml file.

- Similar example for `column_metadata` and `unit_metadata` based on an
existing EML file (just using `eml_read`).

### Automatic Quality Control checks

<!-- Not sure if we'll actually have/need to have all this in place for
this paper?  --> Examples of validation against schema.  Examples checking
md5sums on files or other corruption checks?  Examples checking species
names (taxize?)

### Publish data

Examples of commands for publishing EML files to external repositories.

### Complex search queries

Illustrate how to query for data

### A synthesis platform

Illustrate a basic example of synthesis.  While arbitrary synthesis will
require semantics, a lab could store their data in EML format and build
up there own library of common scripts to extract the data they need. As
a lab database grows, the same queries can continue to deliver more data.
By scripting the data access, analyses can be automatically updated as
the data set grows.

Consider simple example in regression between different morphological
characters of species, or such.

## Challenges

Completely automated data integration will require rich semantic
information. Consider the simple example of combining the two data tables
shown in Figure 2.

<!-- Example illustrating the need for semantics to resolve differences -->

[version 2.1.1]: http://knb.ecoinformatics.org/software/eml/eml-2.1.1
[Temple Lang 2013, XML]: http://cran.r-project.org/web/packages/XML/

### Quality control

The `EML` package performs severak levels of validation on EML files
written from or read into R: validating the XML, and validating against
the EML schema (based on the declared version), and performing additional
checks through the EML online validation tool for elements that cannot
be enforced by schema validation alone.  Validation serves to improve
consistency in data encoding which facilitates the use of EML across a
wide variety of platforms and software.

Several mechanisms are in place to evaluate the quality of the `EML`
package functions themselves.  In addition to the examples from
the documentation and the automated package checks provided for the
R system, the `EML` package includes a complete test suite using the
`testthat` package [Wickham 2011]. Package development on Github includes
continuous integration through the `travis` platform, in which the full
test suite is run every time the code is changed.  This not only gives
users confidence that the development copy can be successfully installed,
but should assist other developers seeking to contribute to the package
to test that their proposed changes have not negatively impacted any of
the existing code base.

### Software reuse and support

The `EML` package is available through the CRAN repository under
a CC0 license. At the time of writing, the current version is `r
packageDescription("EML", fields="Version")`. The current development
branch can be found on Github at [https://github.com/ropensci/EML].
See the NEWS file for changes in recent versions. Users are encouraged
to submit bug reports or feature requests in the issues log.  Contact
`ropensci-discuss@googlegroups.com` or the package maintainer for help.

There are many promising directions for further development of the `EML`
package.  The project's Milestones page provides a list of steps currently
planned or under consideration.  In particular, these milestones include
increased support for interactive mode of entering data (wizards), and
support for integration of semantics defined through the Web Ontology
language (OWL), as discussed in the Challenges section.

### Acknowledgements

This project was supported in part by a grant from the Alfred P Sloan
Foundation (CB and KR), NSF Grant DBI-1306697 (CB), ... `EML` is part
of the rOpenSci project, [http://ropensci.org].