### 2016-03-18 inst/examples/metadata-repositories.Rmd
Draft vignette testing out metadata repositories.

---
title: "Metadata Repositories"
author: "Carl Boettiger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

One of the principle advantages in creating EML is to make it easier to find the data you need, without having to standardize all your data files themselves into one giant database.  Instead, the data files can be whatever you want, provided the relevant information you might search on to discover data of interest is listed in the metadata.

To do this, we will use the `dataone` package to upload a private copy of our data file to the central data repository.

```{r eval = FALSE}
install.packages("dataone", repos= c("https://cran.rstudio.com", "http://nceas.github.io/drat"))
```

```{r}
library("dataone")
library("datapackage")
```

Imagine we have the file paths to our `.csv` data file and an `.xml` EML file providing metadata for it:

```{r}
sampleData <- system.file("extdata/sample.csv", package="dataone")
sampleEML <- system.file("extdata/sample-eml.xml", package="dataone")
```

To upload these to a metadata repository such as [the KNB](http://knb.ecoinformatics.org), we simply create `DataObject`s for both files:

```{r}
dataObj <- new("DataObject", format="text/csv", file=sampleData)
metadataObj <- new("DataObject", format="eml://ecoinformatics.org/eml-2.1.1", file=sampleEML)
```

Note that optionally, `new("DataObject")` could have been given an `id` argument, which could be a (namespaced) UUID from `UUIDgenerate`, or a DOI from a member node (see `generateIdentifier()`).  Since no `id` has been given, a UUID is automatically generated for each.

```{r}
accessRules <- data.frame(subject="CN=Noam Ross A45991,O=Google,C=US,DC=cilogon,DC=org", permission="write")
dataObj <- addAccessRule(dataObj, accessRules)
metadataObj <- addAccessRule(metadataObj, accessRules)
```

We now want to bundle these two objects (data and metadata) into a single "data package" to be uploaded.  To do so, we just create a new `DataPackage` object and then add the data and metadata using the `addData` file:

```{r}
dp <- new("DataPackage")
addData(dp, dataObj, metadataObj)
```

This both adds the files and registers that the metadata object describes the data object.

```{r}
d1c <- D1Client("STAGING", "urn:node:mnStageUCSB2")
packageId <- uploadDataPackage(d1c, dp)
```

Let's see if the ID returned for our package now appears in the DataONE index:

```{r}
query(CNode("STAGING"), searchTerms = list(id = packageId))
```

---------------

Gimme DOI! Note that the example would fail if run here since only the Production (`PROD`) environment can provide DOIs (the `STAGING` environment is only for tests and training examples), and only then on member nodes that offer DOIs, such as the KNB (`urn:node:KNB`).

```{r eval=FLASE}
cn <- CNode("STAGING")
mn <- getMNode(cn, "urn:node:mnStageUCSB2")
newid <- generateIdentifier(mn, "DOI")

## now use the DOI in packaging the DataObject:
new("DataObject", id = newid, format = "eml://ecoinformatics.org/eml-2.1.1", file=sampleEML)
```