### 2018-09-14 old-vignettes/parsing-EML.Rmd
import and nativize eml2  as EML

---
title: "Reading and parsing EML"
author: "Carl Boettiger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parsing EML}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# listviewer

If the `listviewer` package is installed it can provide a convenient way to view and edit EML:

```{r, eval = FALSE}
f <- system.file("tests/eml.xml", package = "emld")
eml <- read_eml(f)
listviewer::jsonedit(eml)
```

# Parsing an EML file

```{r}
f <- system.file("xsd/test/eml-i18n.xml", package = "EML")
eml <- read_eml(f)
```

Here we request all `temporalCoverage` elements occurring in the anywhere in the `eml` document:

```{r}
temporalCoverage <- eml_get(eml, "temporalCoverage")
temporalCoverage
```

Any EML element can be extracted in this way. Let's try an example metadata file for a `dataset` that documents 11 seperate `dataTables`:

```{r}
hf001 <- system.file("examples/hf001.xml", package="EML") 
eml_HARV <- read_eml(hf001)
```

How many `dataTable` entities are there in this dataset?

```{r}
dt <- eml_get(eml_HARV, "dataTable")
length(dt)
```

We can iterate over our list of `dataTable` elements to extract relevant metadata, such as the `entityName` or the download `url`:

```{r}
entities <- eml_get(eml_HARV, "dataTable.entityName")
urls <- sapply(dt, eml_get, "url")
```

Note that the latter example is the same as providing the more verbose arbument that specificies exactly where the `url` of interest is located:

```{r}
urls <- sapply(dt, function(x) x@physical[[1]]@distribution[[1]]@online@url)
```

this verbose syntax can be useful if there are multiple `url` elements in each `dataTable` metadata, and we are trying to get only certain ones and not others. Specifying the exact path in this way can also improve the speed of the command.  For these reasons, programmatic use should consider this format, while the much simpler `eml_get` example shown above is practical for most interactive applications.

Although the default return type for `eml_get` is just the S4 object (whose `print` method displays the corresponding XML structure used to represent that metadata), for a few commonly accessed complex elements, `eml_get` returns a more convenient `data.frame`.  For instance, the `attributeList` describing the metadata for every column in an EML document is returned as a pair of `data.frame`s, one for all the attributes, and an second optional `data.frame` defnining the levels for the factors, if any are used.  Let's take a look:

Here we get the `attributeList` for each `dataTable` in the dataset. We check the length to confirm we get one `attributeList` for each `dataTable`

```{r}
attrs <- eml_get(dt, "attributeList") 
length(attrs)
attrs[[1]]
```

(Note, we could have passed this argument the original `eml_HARV` instead of `dt` here, since we know all `attributeList` elements are inside `dataTable` elements, but this is a bit more explicit and a bit faster.)

This returned `data.frame` object containing the attribute metadata for the first table (hence the `[[1]]`, though `attrs` contains this metadata for all 11 tables now.)  This is the same result we would have gotten using the more explicit call to the helper function `get_attributes()`:

```{r}
get_attributes(eml_HARV@dataset@dataTable[[1]]@attributeList)
```